PANCHANGE_data_preprocessing

### 2020-08-13 PANCHANGE_data_preprocessing2020-07-15.Rmd
Edited the education/employment section to run similarly for each cohort. Added chunks for EDGI and NBR.

# Highest education (prepandemic)
Proceed along qualifications fields. If field is not blank, take that as highest qual. If field is blank, check next field. If all fields are blank, NA
Assumes University degree > A level > NVQ > GCSE/O Level/CSE
```{r Highest education prepandemic}
if(GLAD == TRUE | EDGI == TRUE) {
#create numeric version of the highest education variable
dem.raw.id <- dem.raw.id %>%
  mutate(
    highest_education_numeric =
      case_when(
        demographics.college_or_university_degree_numeric == "1" ~ 4,
        demographics.a_levelsas_levels_or_equivalent_numeric == "1" ~ 3,
        demographics.nvq_or_hnd_or_hnc_or_equivalent_numeric == "1" ~ 2,
        demographics.o_levelsgcses_or_equivalent_numeric == "1" ~ 1,
        demographics.cses_or_equivalent_numeric == "1" ~ 1)
    )

#recode the numeric version into a factor
dem.raw.id <- dem.raw.id %>%
  mutate(
    highest_education =
      recode_factor(
        highest_education_numeric,
        `1` = "GCSE/CSE",
        `2` = "NVQ",
        `3` = "A-levels",
        `4` = "University")
    )

dem.raw.id %>%
  freq(highest_education)
}
```

Check highest education coding
```{r Highest education prepandemic check}
if(GLAD == TRUE | EDGI == TRUE) {
dem.raw.id %>%
  select(highest_education,
        demographics.college_or_university_degree_numeric,
        demographics.a_levelsas_levels_or_equivalent_numeric,
        demographics.nvq_or_hnd_or_hnc_or_equivalent_numeric,
        demographics.o_levelsgcses_or_equivalent_numeric,
        demographics.cses_or_equivalent_numeric
  )
}
```

@@

## Read in data
```{r glad employment cols}
if(GLAD == TRUE){
glad.employment.raw <- readRDS(file = paste0(data.raw_path, "/glad/employment_coping_glad.rds"))
dim(glad.employment.raw)
colnames(glad.employment.raw)
}
```

# Select variables
```{r glad employment cleaning}
if(GLAD == TRUE){
exclude_cols_employment <- c("ID",
                  "Sample")

@@

```{r nbr employment cols}
if(NBR == TRUE){
nbr.employment.raw <- readRDS(file = paste0(data.raw_path, "/nbr/employment_coping_nbr.rds"))
dim(nbr.employment.raw)
colnames(nbr.employment.raw)
}
```

# Select variables
```{r nbr employment cleaning}
if(NBR == TRUE){
exclude_cols_employment <- c("ID",
                  "Sample")

nbr.employment.raw.id <- nbr.employment.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(
    ID = as.numeric(ID)
    ) %>%
  select(
         Sample, # Sample
         ID, # ID
         employment.phd, # level of education - univeristy
         employment.masters_degree_or_equivalent, # level of education - univeristy
         employment.postgraduate_degree_or_equivalent, # level of education - univeristy
         employment.bachelors_degree_or_equivalent, # level of education - univeristy
         employment.a_levelsas_levels_or_equivalent, # level of education - alevel
         employment.o_levelsgcses_or_equivalent, # level of education - gcse
         employment.cses_or_equivalent, # level of education - cse
         employment.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         employment.government_work_key_workers # key worker status
         ) %>%
  add_numeric(., exclude = exclude_cols_employment) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

# Highest education
Proceed along qualifications fields. If field is not blank, take that as highest qual. If field is blank, check next field. If all fields are blank, NA
Assumes University degree > A level > NVQ > GCSE/O Level/CSE
```{r Highest education}
if(GLAD == TRUE | EDGI == TRUE) {
#create numeric version of the highest education variable
glad.employment.raw.id <- glad.employment.raw.id %>%
  mutate(
    highest_education_numeric =
      case_when(
        employment.phd_numeric == "1" ~ 4,
        employment.masters_degree_or_equivalent_numeric == "1" ~ 4,
        employment.postgraduate_degree_or_equivalent_numeric == "1" ~ 4,
        employment.bachelors_degree_or_equivalent_numeric == "1" ~ 4,
        employment.a_levelsas_levels_or_equivalent_numeric == "1" ~ 3,
        employment.nvq_or_hnd_or_hnc_or_equivalent_numeric == "1" ~ 2,
        employment.o_levelsgcses_or_equivalent_numeric == "1" ~ 1,
        employment.cses_or_equivalent_numeric == "1" ~ 1
      )
    )

#recode the numeric version into a factor
glad.employment.raw.id <- glad.employment.raw.id %>%
  mutate(
    highest_education =
      recode_factor(
        highest_education_numeric,
        `1` = "GCSE/CSE",
        `2` = "NVQ",
        `3` = "A-levels",
        `4` = "University")
    )

glad.employment.raw.id %>%
  freq(highest_education)
}
```

Check highest education coding
```{r Highest education check}
if(GLAD == TRUE | EDGI == TRUE) {
glad.employment.raw.id %>%
  select(highest_education,
         employment.phd_numeric,
         employment.masters_degree_or_equivalent_numeric,
         employment.postgraduate_degree_or_equivalent_numeric,
         employment.bachelors_degree_or_equivalent_numeric,
         employment.a_levelsas_levels_or_equivalent_numeric,
         employment.nvq_or_hnd_or_hnc_or_equivalent_numeric,
         employment.o_levelsgcses_or_equivalent_numeric,
         employment.cses_or_equivalent_numeric
  )
}
```

>>>>

# Highest education (prepandemic)
Proceed along qualifications fields. If field is not blank, take that as highest qual. If field is blank, check next field. If all fields are blank, NA
Assumes University degree > A level > NVQ > GCSE/O Level/CSE
```{r Highest education prepandemic}
if(GLAD == TRUE | EDGI == TRUE) {
#create numeric version of the highest education variable
dem.raw.id <- dem.raw.id %>%
  mutate(
    highest_education_prepan_numeric =
      case_when(
        demographics.college_or_university_degree_numeric == "1" ~ 4,
        demographics.a_levelsas_levels_or_equivalent_numeric == "1" ~ 3,
        demographics.nvq_or_hnd_or_hnc_or_equivalent_numeric == "1" ~ 2,
        demographics.o_levelsgcses_or_equivalent_numeric == "1" ~ 1,
        demographics.cses_or_equivalent_numeric == "1" ~ 1)
    )

#recode the numeric version into a factor
dem.raw.id <- dem.raw.id %>%
  mutate(
    highest_education_prepan =
      recode_factor(
        highest_education_numeric,
        `1` = "GCSE/CSE",
        `2` = "NVQ",
        `3` = "A-levels",
        `4` = "University")
    )

dem.raw.id %>%
  freq(highest_education_prepan)
}
```

Check highest education coding
```{r Highest education prepandemic check}
if(GLAD == TRUE | EDGI == TRUE) {
dem.raw.id %>%
  select(highest_education_prepan,
        demographics.college_or_university_degree_numeric,
        demographics.a_levelsas_levels_or_equivalent_numeric,
        demographics.nvq_or_hnd_or_hnc_or_equivalent_numeric,
        demographics.o_levelsgcses_or_equivalent_numeric,
        demographics.cses_or_equivalent_numeric
  )
}
```

@@

## GLAD
### Read in data
```{r glad employment cols}
if(GLAD == TRUE){
glad.employment.raw <- readRDS(file = paste0(data.raw_path, "/glad/employment_coping_glad.rds"))
dim(glad.employment.raw)
colnames(glad.employment.raw)
}
```

### Select variables
```{r glad employment cleaning}
if(GLAD == TRUE){
exclude_cols_employment <- c("ID",
                  "Sample")

@@

```{r glad copy employment}
if(GLAD == TRUE){
  employment.raw.id <- glad.employment.raw.id
}
```

## EDGI
### Read in data
```{r edgi employment cols}
if(EDGI == TRUE){
edgi.employment.raw <- readRDS(file = paste0(data.raw_path, "/edgi/employment_coping_edgi.rds"))
dim(edgi.employment.raw)
colnames(edgi.employment.raw)
}
```

### Select variables
```{r edgi employment cleaning}
if(EDGI == TRUE){
exclude_cols_employment <- c("ID",
                  "Sample")

edgi.employment.raw.id <- edgi.employment.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(
    ID = as.numeric(ID)
    ) %>%
  select(
         Sample, # Sample
         ID, # ID
         employment.phd, # level of education - univeristy
         employment.masters_degree_or_equivalent, # level of education - univeristy
         employment.postgraduate_degree_or_equivalent, # level of education - univeristy
         employment.bachelors_degree_or_equivalent, # level of education - univeristy
         employment.a_levelsas_levels_or_equivalent, # level of education - alevel
         employment.o_levelsgcses_or_equivalent, # level of education - gcse
         employment.cses_or_equivalent, # level of education - cse
         employment.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         employment.government_work_key_workers # key worker status
         ) %>%
  add_numeric(., exclude = exclude_cols_employment) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(edgi.employment.raw.id)
# Inspect colnames
colnames(edgi.employment.raw.id)
#Differences
dim(edgi.employment.raw)[1]-dim(edgi.employment.raw.id)[1]
}
```

```{r edgi copy employment}
if(EDGI == TRUE){
  employment.raw.id <- edgi.employment.raw.id
}
```

## NBR
### Read in data
```{r nbr employment cols}
if(NBR == TRUE){
nbr.employment.raw <- readRDS(file = paste0(data.raw_path, "/nbr/employment_coping_nbr.rds"))
dim(nbr.employment.raw)
colnames(nbr.employment.raw)
}
```

### Select variables
```{r nbr employment cleaning}
if(NBR == TRUE){
exclude_cols_employment <- c("ID",
                  "Sample")

nbr.employment.raw.id <- nbr.employment.raw %>%
  drop_na(subjectid) %>% # Drop NAs
  distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(subjectid, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(
    ID = as.numeric(ID)
    ) %>%
  select(
         Sample, # Sample
         ID, # ID
         employment.phd, # level of education - univeristy
         employment.masters_degree_or_equivalent, # level of education - univeristy
         employment.postgraduate_degree_or_equivalent, # level of education - univeristy
         employment.bachelors_degree_or_equivalent, # level of education - univeristy
         employment.a_levelsas_levels_or_equivalent, # level of education - alevel
         employment.o_levelsgcses_or_equivalent, # level of education - gcse
         employment.cses_or_equivalent, # level of education - cse
         employment.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         employment.government_work_key_workers # key worker status
         ) %>%
  add_numeric(., exclude = exclude_cols_employment) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

```{r nbr copy employment}
if(NBR == TRUE){
  employment.raw.id <- nbr.employment.raw.id
}
```

## Highest education (baseline)
Proceed along qualifications fields. If field is not blank, take that as highest qual. If field is blank, check next field. If all fields are blank, NA
Assumes University degree > A level > NVQ > GCSE/O Level/CSE
```{r Highest education baseline}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE) {
#create numeric version of the highest education variable
employment.raw.id <- employment.raw.id %>%
  mutate(
    highest_education_numeric =
      case_when(
        employment.phd_numeric == "1" ~ 4,
        employment.masters_degree_or_equivalent_numeric == "1" ~ 4,
        employment.postgraduate_degree_or_equivalent_numeric == "1" ~ 4,
        employment.bachelors_degree_or_equivalent_numeric == "1" ~ 4,
        employment.a_levelsas_levels_or_equivalent_numeric == "1" ~ 3,
        employment.nvq_or_hnd_or_hnc_or_equivalent_numeric == "1" ~ 2,
        employment.o_levelsgcses_or_equivalent_numeric == "1" ~ 1,
        employment.cses_or_equivalent_numeric == "1" ~ 1
      )
    )

#recode the numeric version into a factor
employment.raw.id <- employment.raw.id %>%
  mutate(
    highest_education =
      recode_factor(
        highest_education_numeric,
        `1` = "GCSE/CSE",
        `2` = "NVQ",
        `3` = "A-levels",
        `4` = "University")
    )

employment.raw.id %>%
  freq(highest_education)
}
```

Check highest education coding
```{r Highest education check}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE) {
employment.raw.id %>%
  select(highest_education,
         employment.phd_numeric,
         employment.masters_degree_or_equivalent_numeric,
         employment.postgraduate_degree_or_equivalent_numeric,
         employment.bachelors_degree_or_equivalent_numeric,
         employment.a_levelsas_levels_or_equivalent_numeric,
         employment.nvq_or_hnd_or_hnc_or_equivalent_numeric,
         employment.o_levelsgcses_or_equivalent_numeric,
         employment.cses_or_equivalent_numeric
  )
}
```
